---
title: "Project 4"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(here)
library(readr)
library(dplyr)
library(stringr)
library(purrr)
library(ggplot2)
library(omnibus)

# test if a directory named "data" exists locally
if(!dir.exists(here("data"))) { dir.create(here("data")) }

if(!file.exists(here("data", "tuition_cost.RData")) | !file.exists(here("data", "salary_potential.RData"))){
  tuition_cost <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_cost.csv')
  salary_potential <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/salary_potential.csv')
  
  # save the files to RData objects
  save(tuition_cost, file = here("data", "tuition_cost.RData"))
  save(salary_potential, file = here("data", "salary_potential.RData"))
}

# read in the data locally
tuition_cost <- get(load(here("data", "tuition_cost.RData")))
salary_potential <- get(load(here("data", "salary_potential.RData")))
```

About
=======================================================================
### Purpose of the dashboard

### Original location of the data
- The data originally came from the US Department of Education. I found the data on [TidyTuesday](https://www.tidytuesday.com) and the link to its GitHub repository is <https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-03-10>.

Column {data-width=650}
-----------------------------------------------------------------------

### Chart A

```{r}

```

Column {data-width=350}
-----------------------------------------------------------------------

### Chart B

```{r}

```

### Chart C

```{r}

```

Data
=======================================================================
- There are 2 datasets: tuition_cost which contains the in-state and out-state tuition for each school and salary_potential which contains early and mid career pay in USD for students in each school.
- Data dictionaries are available here:
<https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-03-10#tuition_costcsv>,
<https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-03-10#salary_potentialcsv>.


Static 1
=======================================================================


Static 2
=======================================================================

Interactive 1
=======================================================================


Interactive 2
=======================================================================


Analysis
=======================================================================
### Data Wrangling
```{r}
# We first combine the 2 data frames. We include only schools that are in both data frames.
combined <- inner_join(tuition_cost, salary_potential, by = "name")
```

```{r}
# Remove unwanted variables
combined <- combined %>% select(-room_and_board, -in_state_total, -out_of_state_total, -rank, -make_world_better_percent)

# Check if there's any NAs. If missing values exist, get the corresponding row numbers.
naRows(combined)
```

Luckily, we don't have any missing values.

```{r}
# We want to check whether there's any discrepancy between the "state" variables from the 2 data frames.
temp <- map2(.x = combined$state, .y = combined$state_name, .f = ~str_equal(.x, .y))

# Save the index of not matched cases
ind <- which(temp == FALSE)

combined %>% slice(ind) %>% select(state, state_name)
```

From here, we see that it is possible that the difference in the 2 columns is caused by the dash in the values of "state_name". We're going to check that.

```{r}
# Replace "-" with " "
combined$state_name <- str_replace(combined$state_name, "-", " ")
temp2 <- pmap(list(x = combined$state, y = combined$state_name),
			  .f = ~str_equal(..1, ..2))
ind2 <- which(temp2 == FALSE)
combined %>% slice(ind2) %>% select(name, state, state_code, state_name)
```

We remove the 4 observations with inconsistent information.
```{r}
combined <- combined %>% slice(-ind2) %>% select(-state_name)
combined
```

```{r}
combined %>% count(degree_length)
```

We keep only the 4-year schools.
```{r}
combined <- combined %>% filter(degree_length == "4 Year")
combined
```

We want to find the modal value(s) for the percent of student body in STEM:
```{r}
mmode(combined$stem_percent)
```

We can see that the modes are 0 and 7.

### Data Visualization
- Plot 1
```{r}
combined %>% ggplot(aes(x = type, y = out_of_state_tuition)) +
	geom_boxplot() + 
	labs(title = "Out-of-state Tuition by School Type", 
		 subtitle = "This is a box plot displaying the summary statistics for the out-of-state tuition by school type.\nWe can see that the first quartile of private school tuition is above the third quartile of the\npublic school tuition.", 
		 caption = "Mincen Liu") +
	xlab("school type") + ylab("out-of-state tuition in USD")
```

- Plot 2
```{r}
combined %>% ggplot(aes(x = out_of_state_tuition, y = mid_career_pay)) + 
	geom_point() + 
	facet_grid(type ~ .) + 
	labs(title = "Mid Career Pay vs. Out-of-state Tuition by School Type", 
		 subtitle = "This is a scatter plot showing the relationship between the mid career pay and the out-of-state\ntuition, faceted by school type. We see an overall positive relation in each school type.", 
		 caption = "Mincen Liu") +
	xlab("out-of-state tuition in USD") + ylab("mid career pay in USD")
```

- Plot 3
```{r}
combined %>% ggplot(aes(early_career_pay)) +
	geom_density() +
	labs(title = "Early Career Pay", 
		 subtitle = "This is a density plot of the early career pay. We see a right-skewed distribution.", 
		 caption = "Mincen Liu") +
	xlab("early career pay in USD") + ylab("density")
```

### Interpretation
- On average, private school out-of-state tuition is much higher than public school out-of-state tuition among the 4-year institutions.
- We see an overall positive relation between the mid career pay and the out-of-state tuition in each school type.
- The early career pay has a right-skewed distribution.

### List of Functions Used
- `dplyr`: `glimpse()`, `inner_join()`, `select()`, `slice()`, `count()`, `filter()`
- `stringr`: `str_equal()`, `str_replace()`
- `purrr`: `map2()`, `pmap()`
- `ggplot2`: `geom_boxplot()`, `geom_point()`, `geom_density()`, `facet_grid()`
- `omnibus`: `naRows()`, `mmode()`

