---
title: "Project 4"
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(here)
library(readr)
library(dplyr)
library(stringr)
library(purrr)
library(ggplot2)
library(omnibus)
library(DT)
library(htmltools)

# test if a directory named "data" exists locally
if(!dir.exists(here("data"))) { dir.create(here("data")) }

if(!file.exists(here("data", "tuition_cost.RData")) | !file.exists(here("data", "salary_potential.RData"))){
  tuition_cost <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_cost.csv')
  salary_potential <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/salary_potential.csv')
  
  # save the files to RData objects
  save(tuition_cost, file = here("data", "tuition_cost.RData"))
  save(salary_potential, file = here("data", "salary_potential.RData"))
}

# read in the data locally
tuition_cost <- get(load(here("data", "tuition_cost.RData")))
salary_potential <- get(load(here("data", "salary_potential.RData")))
```

About {data-icon="fa-question-circle"}
=======================================================================

**What is the purpose of this dashboard?**

This dashboard is to illustrate the college tuition, the pay after graduation, and their relationship.

**The data**

The data used in this dashboard originally came from the US Department of Education. I found the data on [TidyTuesday](https://www.tidytuesday.com) and the link to its GitHub repository is <https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-03-10>.


Data {data-icon="fa-database"}
=======================================================================

Column {data-width = 500}
-------------------------------------
**tuition_cost** contains the in-state and out-of-state tuition for each school.

Data dictionary: <https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-03-10#tuition_costcsv>

```{r, echo=FALSE}
tc_table <- tuition_cost %>% select(name, state, type, in_state_tuition, out_of_state_tuition)
```

```{r, echo=FALSE}
renderDT({
  DT::datatable(tc_table, caption = tags$caption(style = 'caption-side: top; text-align: Left;'), 
            options = list(autowidth = TRUE, pageLength = 15, searching = FALSE, scroller = TRUE, scrollY = '500px'))
})
```

Column {data-width = 500}
-------------------------------------
**salary_potential** contains the early and mid career pay in USD for students in each school.

Data dictionary: <https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-03-10#salary_potentialcsv>

```{r, echo=FALSE}
sp_table <- salary_potential %>% rename(state = state_name) %>%
  select(name, state, early_career_pay, mid_career_pay, stem_percent)
```

```{r, echo=FALSE}
renderDT({
  DT::datatable(sp_table, caption = tags$caption(style = 'caption-side: top; text-align: Left;'), 
            options = list(autowidth = TRUE, pageLength = 15, searching = FALSE, scroller = TRUE, scrollY = '500px'))
})
```


School Counts {data-icon="fa-school"}
=======================================================================

Column {.tabset .tabset-fade}
-------------------------------------

### Total Schools Per State
```{r, echo=FALSE}
tuition_cost %>% ggplot(aes(x = state_code)) + 
  geom_bar() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 6)) +
  xlab("state") + labs(title = "Total Number of Schools Per State")
```

### Public Schools Per State
```{r, echo=FALSE}
tuition_cost %>% filter(type == "Public") %>%
  ggplot(aes(x = state_code)) + 
  geom_bar() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 6)) +
  xlab("state") + labs(title = "Total Number of Public Schools Per State")
```

### Private Schools Per State
```{r, echo=FALSE}
tuition_cost %>% filter(type == "Private") %>%
  ggplot(aes(x = state_code)) + 
  geom_bar() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 6)) +
  xlab("state") + labs(title = "Total Number of Private Schools Per State")
```


Tuition vs. Pay {data-icon="fa-school"}
=======================================================================

### **Average in-state tuition**

```{r, echo=FALSE}
valueBox(value = formatC(mean(tuition_cost$in_state_tuition), 7),
         color = "white")
```

### **Average out-of-state tuition**

```{r, echo=FALSE}
valueBox(value = formatC(mean(tuition_cost$out_of_state_tuition), 7),
         color = "white")
```

### **Average early career pay**

```{r, echo=FALSE}
valueBox(value = formatC(mean(salary_potential$early_career_pay), 7),
         color = "white")
```

### **Average mid career pay**

```{r, echo=FALSE}
valueBox(value = formatC(mean(salary_potential$mid_career_pay), 7),
         color = "white")
```


Mid Career Pay {data-icon="fa-seedling"}
=======================================================================

Column {.sidebar data-width=300}
-----------------------------------------------------------------------

```{r, echo=FALSE}
selectInput(inputId = "state_selected",
              label = "Select a state to explore:",
            choices = salary_potential %>%
              pull(state_name) %>%
              unique() %>%
              sort(), selected = "Alabama")
```

Column {data-width=500}
-----------------------------------------------------------------------

### Top 10 schools with the highest mid career pay in the selected state

```{r, echo=FALSE}
renderPlot({
  dat <- salary_potential %>% filter(state_name == input$state_selected) %>% arrange(desc(mid_career_pay)) %>% slice_head(n = 10)

  dat %>% ggplot(aes(x = mid_career_pay , y = reorder(name, mid_career_pay))) + geom_col() + 
    ylab("name of school") + xlab("estimated mid career pay in USD")
})
```


Stem {data-icon="fa-seedling"}
=======================================================================

Column {.sidebar data-width=300}
-----------------------------------------------------------------------

```{r, echo=FALSE}
numericInput("perc_stem", label = "Percent of student body in STEM:",
             min = 0, max = 100, value = 20, step = 1)
```

Column {data-width=500}
-----------------------------------------------------------------------

### Schools with input percent of student body in STEM:

```{r}
renderDT({
  dat2 <- salary_potential %>% filter(stem_percent == input$perc_stem) %>% select(name, early_career_pay, mid_career_pay)
  
  datatable(dat2)
})
```


Analysis
=======================================================================

Column {.tabset}
-------------------------------------

### Data Wrangling

```{r, echo=TRUE}
# We first combine the 2 data frames. We include only schools that are in both data frames.
combined <- inner_join(tuition_cost, salary_potential, by = "name")
```

```{r, echo=TRUE}
# Remove unwanted variables
combined <- combined %>% select(-room_and_board, -in_state_total, -out_of_state_total, -rank, -make_world_better_percent)

# Check if there's any NAs. If missing values exist, get the corresponding row numbers.
naRows(combined)
```

Luckily, we don't have any missing values.

```{r, echo=TRUE}
# We want to check whether there's any discrepancy between the "state" variables from the 2 data frames.
temp <- map2(.x = combined$state, .y = combined$state_name, .f = ~str_equal(.x, .y))

# Save the index of not matched cases
ind <- which(temp == FALSE)

combined %>% slice(ind) %>% select(state, state_name)
```

From here, we see that it is possible that the difference in the 2 columns is caused by the dash in the values of "state_name". We're going to check that.

```{r, echo=TRUE}
# Replace "-" with " "
combined$state_name <- str_replace(combined$state_name, "-", " ")
temp2 <- pmap(list(x = combined$state, y = combined$state_name),
			  .f = ~str_equal(..1, ..2))
ind2 <- which(temp2 == FALSE)
combined %>% slice(ind2) %>% select(name, state, state_code, state_name)
```

We remove the 4 observations with inconsistent information.
```{r, echo=TRUE}
combined <- combined %>% slice(-ind2) %>% select(-state_name)
combined
```

```{r, echo=TRUE}
combined %>% count(degree_length)
```

We keep only the 4-year schools.
```{r, echo=TRUE}
combined <- combined %>% filter(degree_length == "4 Year")
combined
```

We want to find the modal value(s) for the percent of student body in STEM:
```{r, echo=TRUE}
mmode(combined$stem_percent)
```

We can see that the modes are 0 and 7.

### Data Visualization {.tabset}

#### Plot 1
```{r, echo=TRUE}
combined %>% ggplot(aes(x = type, y = out_of_state_tuition)) +
	geom_boxplot() + 
	labs(title = "Out-of-state Tuition by School Type", 
		 subtitle = "This is a box plot displaying the summary statistics for the out-of-state tuition by school type.\nWe can see that the first quartile of private school tuition is above the third quartile of the\npublic school tuition.", 
		 caption = "Mincen Liu") +
	xlab("school type") + ylab("out-of-state tuition in USD")
```

#### Plot 2
```{r, echo=TRUE}
combined %>% ggplot(aes(x = out_of_state_tuition, y = mid_career_pay)) + 
	geom_point() + 
	facet_grid(type ~ .) + 
	labs(title = "Mid Career Pay vs. Out-of-state Tuition by School Type", 
		 subtitle = "This is a scatter plot showing the relationship between the mid career pay and the out-of-state\ntuition, faceted by school type. We see an overall positive relation in each school type.", 
		 caption = "Mincen Liu") +
	xlab("out-of-state tuition in USD") + ylab("mid career pay in USD")
```

#### Plot 3
```{r, echo=TRUE}
combined %>% ggplot(aes(early_career_pay)) +
	geom_density() +
	labs(title = "Early Career Pay", 
		 subtitle = "This is a density plot of the early career pay. We see a right-skewed distribution.", 
		 caption = "Mincen Liu") +
	xlab("early career pay in USD") + ylab("density")
```

### Interpretation
- On average, private school out-of-state tuition is much higher than public school out-of-state tuition among the 4-year institutions.
- We see an overall positive relation between the mid career pay and the out-of-state tuition in each school type.
- The early career pay has a right-skewed distribution.
