---
title: "Untitled"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(here)
library(readr)

# test if a directory named "data" exists locally
if(!dir.exists(here("data"))) { dir.create(here("data")) }

if(!file.exists(here("data", "tuition_cost.RData")) | !file.exists(here("data", "salary_potential.RData"))){
  tuition_cost <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_cost.csv')
  salary_potential <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/salary_potential.csv')
  
  # save the files to RData objects
  save(tuition_cost, file = here("data", "tuition_cost.RData"))
  save(salary_potential, file = here("data", "salary_potential.RData"))
}

# read in the data locally
tuition_cost <- get(load(here("data", "tuition_cost.RData")))
salary_potential <- get(load(here("data", "salary_potential.RData")))
```

About
=======================================================================

Column {data-width=650}
-----------------------------------------------------------------------

### Chart A

```{r}

```

Column {data-width=350}
-----------------------------------------------------------------------

### Chart B

```{r}

```

### Chart C

```{r}

```

Data
=======================================================================
- The data originally came from the US Department of Education. I found the data on [TidyTuesday](https://www.tidytuesday.com) and the link to its GitHub repository is <https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-03-10>.
- I use 2 datasets from this repository: tuition_cost.csv which contains the in-state and out-state tuition for each school and salary_potential.csv which contains early and mid career pay in USD for students in each school.
- Data dictionaries are available here:
<https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-03-10#tuition_costcsv>,
<https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-03-10#salary_potentialcsv>.





Analysis
=======================================================================